Digging Deeper Into the State of the Practice for Domain Specific Research Software

### 2022-02-17 StateOfPractice/Papers/ICCS-2022-SOP_MethodologyPlusExamples/DigDeep.tex
Additional row for Table 3

\begin{table}[!h]
\begin{center}
\begin{tabular}{ p{3cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm} }
\toprule
~ \ & \cite{USGS2019} & \cite{TobiasEtAl2018} & \cite{BrettEtAl2021} & \cite{WilsonEtAl2016} & \cite{SmithAndRoscoe2018} & \cite{HerouxEtAl2008} & \cite{ThielEtAl2020} & \cite{vanGompelEtAl2016} & \cite{OrvizEtAl2017}\\
\midrule
LICENSE & \checkmark &  & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark\\
README &  &  &  &  & & & & \\
CONTRIBUTING &  &  &  &  & & & & & \\
CITATION &  &  &  &  & & & & & \\
CHANGELOG &  &  &  &  & & & & & \\
INSTALL &  &  &  &  & & & & & \\
Uninstall &  &  &  &  & & & & & \\
Contributor Guide &  &  &  &  & & & & & \\
Getting started &  &  &  &  & & & & & \\
User manual &  &  &  &  & & & & & \\
Tutorials &  &  &  &  & & & & & \\
FAQ &  &  &  &  & & & & & \\
Dependency List &  &  &  & & & & & & \\
Issue Track &  &  &  &  & & & & & \\
Version Control &  &  &  &  & & & & &\\ 
Requirements &  &  &  &  & & & & & \\
Design &  &  &  &  & & & & & \\
API Doc. &  &  &  &  & & & & & \\
Build Scripts &  &  &  &  & & & & & \\
Unit Tests &  &  &  &  & & & & & \\
Test Plan &  &  &  &  & & & & & \\
Integ. Tests &  &  &  &  & & & & &  \\
System Tests &  &  &  &  & & & & & \\
Acceptance Tests &  &  &  &  & & & & &  \\
Regression Tests &  &  &  &  & & & & & \\
Code Style Guide &  &  &  &  & & & & & \\
Release Info. &  &  &  &  & & & & & \\
Product Roadmap &  &  &  &  & & & & & \\
\midrule
\end{tabular}
\caption{Commonly Recommended Artifacts in Software Development Guidelines} \label{Tbl_Guidelines}
\end{center}
\end{table}

Another point of comparison is to look at the use of requirements related
artifacts.  Software requirements documentation provides a good example of the
common deviation between recommendations and practice.  Although requirements
documentation is recommended by some
\cite{TobiasEtAl2018,HerouxEtAl2008,SmithAndKoothoor2016}, in practice
scientific software developers often do not produce a proper requirements
specification \cite{HeatonAndCarver2015}. Sanders and Kelly \cite{SandersAndKelly2008} interviewed
16 scientists from 10 disciplines and found that none of the scientists created
requirements specifications, unless regulations in their field mandated such a
document. \cite{Nguyen-HoanEtAl2010} showed requirements are the least commonly
produced type of documentation for scientific software in general. When looking
at the pain points for research software developers, \cite{WieseEtAl2019} found
that software requirements and management is the software engineering discipline
that most hurts scientific developers, accounting for 23 \% of the technical
problems reported by study participants.  The lack of support for requirements
is likely due to the perception that up-front requirements are impossible for
research software \cite{CarverEtAl2007,SegalAndMorris2008}, but when the
instance on ``up-front'' requirements is dropped, allowing the requirements to
be written iteratively and incrementally, requirements are feasible
\cite{Smith2016}.

The results for the LBM example are shown in Table~\ref{artifactspresent}.  The
majority of LBM generated artifacts correspond to general recommendations from
research software developers.  A union of the three columns in
Table~\ref{artifactspresent} mostly corresponds to recommendations made by the
research software community, as shown in Table~\ref{}.

@@

Although the LBM community participates in most of the practices we found listed
in the general research software guidelines, some recommended practices were not
observed or rarely observed. For instance API documentation was rarely observed
for LBM software, but it is a frequently recommended artifact
\cite{SmithAndRoscoe2018,ThielEtAl2020,vanGompelEtAl2016,OrvizEtAl2017,SSI2022,Zadka2018}.
In addition to the items in the last column of Table~\ref{artifactspresent}, we
can add the following community recommended items that we rarely, if ever,
observed:

\begin{itemize}
\item A roadmap showing what is planned for the future
\cite{Yo2021,ThielEtAl2020,vanGompelEtAl2016}.  We did see this information
mentioned in some repos, but we did not observe a specific artifact devoted to
this purpose.
\item A code of conduct to explicitly say how developers should treat one
another \cite{Yo2021,ThielEtAl2020}.  We did not observe a policy of this
sort, possibly because the number of developers for each LBM project is small.
\item Programming style guidelines so that programming labels and formatting are
consistent \cite{TobiasEtAl2018,ThielEtAl2020,OrvizEtAl2017,Zadka2018,vanGompelEtAl2016,WilsonEtAl2014}.  We saw this information as part of
some developer guides, but only rarely.
\item Checklists can be used in projects to ensure that best practices are
followed by all developers.  Some examples include checklists merging branches
into master \cite{Brown2015}, checklists for saving and sharing changes to the
project \cite{WilsonEtAl2016}, checklists for new and departing team members
\cite{HerouxAndBernholdt2018}, checklists for processes related to commits and
releases \cite{HerouxEtAl2008} and checklists for overall software quality
\cite{ThielEtAl2020,SSI2022}.  For LBM software, ESPResSo has a checklist for
managing releases, but otherwise they are used only rarely for LBM.
\item Uninstall instructions \cite{vanGompelEtAl2016} were not observed for any
of the LBM projects.
\end{itemize}

>>>>

\begin{table}[!h]
\begin{center}
\begin{tabular}{ p{3cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm}p{1cm} }
\toprule
~ \ & \cite{USGS2019} & \cite{TobiasEtAl2018} & \cite{BrettEtAl2021} & \cite{WilsonEtAl2016} & \cite{SmithAndRoscoe2018} & \cite{HerouxEtAl2008} & \cite{ThielEtAl2020} & \cite{vanGompelEtAl2016} & \cite{OrvizEtAl2017}\\
\midrule
LICENSE & \checkmark &  & \checkmark & \checkmark & \checkmark & & \checkmark & \checkmark & \checkmark\\
README &  &  &  &  & & & & \\
CONTRIBUTING &  &  &  &  & & & & & \\
CITATION &  &  &  &  & & & & & \\
CHANGELOG &  &  &  &  & & & & & \\
INSTALL &  &  &  &  & & & & & \\
Uninstall &  &  &  &  & & & & & \\
Contributor Guide &  &  &  &  & & & & & \\
Getting started &  &  &  &  & & & & & \\
User manual &  &  &  &  & & & & & \\
Tutorials &  &  &  &  & & & & & \\
FAQ &  &  &  &  & & & & & \\
Dependency List &  &  &  & & & & & & \\
Issue Track &  &  &  &  & & & & & \\
Version Control &  &  &  &  & & & & &\\ 
Requirements &  &  &  &  & & & & & \\
Design &  &  &  &  & & & & & \\
API Doc. &  &  &  &  & & & & & \\
Build Scripts &  &  &  &  & & & & & \\
Unit Tests &  &  &  &  & & & & & \\
Test Plan &  &  &  &  & & & & & \\
Integ. Tests &  &  &  &  & & & & &  \\
System Tests &  &  &  &  & & & & & \\
Acceptance Tests &  &  &  &  & & & & &  \\
Regression Tests &  &  &  &  & & & & & \\
Code Style Guide &  &  &  &  & & & & & \\
Release Info. &  &  &  &  & & & & & \\
Product Roadmap &  &  &  &  & & & & & \\
Code of Conduct &  &  &  &  & & & & & \\
\midrule
\end{tabular}
\caption{Commonly Recommended Artifacts in Software Development Guidelines} \label{Tbl_Guidelines}
\end{center}
\end{table}

The results for the LBM example are shown in Table~\ref{artifactspresent}.  The
majority of LBM generated artifacts correspond to general recommendations from
research software developers.  A union of the three columns in
Table~\ref{artifactspresent} mostly corresponds to recommendations made by the
research software community, as shown in Table~\ref{Tbl_Guidelines}.

@@

Although the LBM community participates in most of the practices we found listed
in the general research software guidelines, some recommended practices were not
observed or rarely observed. For instance API documentation was rarely observed
for LBM software, but it is a frequently recommended (Table~\ref{Tbl_Guidelines}).
In addition to the items in the last column of Table~\ref{artifactspresent}, we
can add the following community recommended items that we rarely, if ever,
observed:

\begin{itemize}
\item A roadmap showing what is planned for the future.
\item A code of conduct to explicitly say how developers should treat one
another.
\item Programming style guidelines so that programming labels and formatting are
consistent \cite{TobiasEtAl2018,ThielEtAl2020,OrvizEtAl2017,Zadka2018,vanGompelEtAl2016,WilsonEtAl2014}.  We saw this information as part of
some developer guides, but only rarely.
\item Checklists can be used in projects to ensure that best practices are
followed by all developers.  Some examples include checklists merging branches
into master \cite{Brown2015}, checklists for saving and sharing changes to the
project \cite{WilsonEtAl2016}, checklists for new and departing team members
\cite{HerouxAndBernholdt2018}, checklists for processes related to commits and
releases \cite{HerouxEtAl2008} and checklists for overall software quality
\cite{ThielEtAl2020,SSI2022}.  For LBM software, ESPResSo has a checklist for
managing releases, but otherwise they are used only rarely for LBM.
\item Uninstall instructions \cite{vanGompelEtAl2016} were not observed for any
of the LBM projects.
\end{itemize}
